Supporting information: Tail associations and their impact on extinction risk

### 2019-10-28 MainText.Rmd
working on para intro-ing partial spearman

<!--concept for correlation in the extreme tail: tail association-->
Though earlier literature considered the influence of spatial synchrony 
in environmental variables on metapopulation extinction risk, to 
our knowledge the distinct influence of "asymmetric tail associations" between
environmental time series in different locations has hardly been considered. 
"Tail associations" are
associations between two random quantities in their extreme values. 
For instance, if two positively associated 
variables are more closely associated in their smaller values than in their larger values,
they are said to have stronger
"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). 
If two positively associated variables are more associated in their larger 
values than in their smaller values, they are 
said to have stronger "right-tail" or "upper-tail" association 
(Figure \ref{fig_intro_cop}C, D). 
Figure \ref{fig_intro_cop} shows that two variables can have a wide 
range of patterns of tail association, from extreme left- to extreme right-tail
association, with the same overall correlation. So tail association
is a distinct concept from correlation itself, and may have distinct ecological
ramifications [@Ghosh_copula]. Asymmetric tail associations have been documented 
between environmental variables, and between the same environmental variable 
measured in different places [@Serinaldi2008; @li2013; @She2018; @Goswami2018]. 
Asymmetric tail associations have also been documented between numerous ecological
variables and between environmental and ecological variables [@Ghosh_copula].

@@

There are also reasons to hypothesize that asymmetry of tail association in 
environmental variables may influence extinction risks for metapopulations.
For instance, assuming for the sake of a simple example that low values of an environmental
variable are "bad" for populations of a given species and high values are "good",
greater left-tail associations between measurements of the variable in different
habitat patches may cause higher metapopulation extinction risk because then
very bad years for component populations would occur at the same time in 
many patches, reducing the potential for rescue effects. Greater left-tail associations
would probably not have the same effect, even if overall correlations between locations
were the same, because in that case it would be very good years for the component 
populations that would occur at the same time in many patches [@Ghosh_copula]. 
@Ghosh_copula substantiated this intuition by showing that it held true for a 
modeling setup using a spatial version of a very simple non-density dependent
population growth model (the Lewontin-Cohen model). In essence, if environmental
"catastrophes" (i.e., extremely bad years for a population) are widely 
spatially synchronized it should create much greater metapopulation extinction 
risk than if "bonanzas" (i.e., extremely good years) are widely synchronized, even
if overall (non-tail-specific) levels of environmental synchrony are the same
in both these cases. The present study extends tests of this idea to metapopulation
models incorporating density dependence.

@@

<!--objectives, last para of intro-->
In this paper, we will explore the potential for asymmetric tail associations to 
influence extinction risk in populations and metapopulations. Specifically,
we will address the following sets of questions. Q1) Can the nature of tail associations
between vital rates in an age-structured population model influence
extinction risk for the model, and to what extent? Q2) Does the nature
of tail associations between measurements of an environmental variable in different
locations influence the extinction risk of a density-dependent metapopulation model
that depends on the environmental variable? How does the nature of the density
dependence mediate this influence? Will density dependence reverse or reproduce
the earlier result [@Ghosh_copula], using non-density-dependent models, that 
left-tail dependence in environmental variables accentuates and right-tail dependence 
mitigates extinction risks?
For Q1, we simulate a simple 2-stage model. We hypothesize that left-tail dependence
between juvenile survival and adult fecundity rates increases extinction risk
relative to right-tail dependence, even when the overall correlation is kept fixed. 
For Q2, we simulate spatial extensions of several simple and commonly
used population models. We intentionally use a simulation approach and simple models
instead of attampting mathematical analysis and/or using realistic models
because our intention here is to provide an initial exploration only of 
the potential mechanisms outlined above which suggest that
asymmetries of tail association may make a difference for extinction risks. 
<!--***DAN: In Discussion need to talk about what analytic results are needed
to take the next step of determining when tail depencence may make enough of
a difference to think about. I think our results, together with Tulja's 
approximation, suggest that tail associations probably don't make much difference
for models without a spatial component. But our results illsutaret that it can
make a big difference when there is a spatial component, apparently because of 
the influence of tail associations on rescue effects. Could maybe even outline
analytic approaches for non-dd models? Not sure what would be done with dd
models. The point is you need to take the next step in the Discussion of 
talking about how our results, which admitedly are just for specific models,
could be turned into analytic results or general results.-->
Though tail associations are used
to assess risks in fields such as finance and hydrology 
[@alexander2008market; @li2013; @chen2011flood; @Goswami2018], to our
knowledge asymmetric tail associations have rarely been assessed for their
importance for extinction risks of populations and species (but see @Ghosh_copula). 
Our questions are an important step for understanding
the potential influence of ECEs on extinction risks if ECEs are associated with
tail dependence in environmental variables.

@@

<!--FIG : pedagogical fig showing 4 types of tail associations: 
extreme left, moderate left, moderate right, extreme right-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
\caption{Two stochastic quantities can have same Kendall 
correlation, but different degrees of tail association: 
(A) extreme left-tail association, (B) moderate left-tail association, (C) 
moderate right-tail association, and (D) extreme right-tail association.
In all cases, Kendall's $\tau$ was the same, $\tau = 0.75$,
up to sample variation (the sample correlation for each dataset is displayed
on its panel). The extreme left-tail (respectively, right-tail) dependent 
case is perfectly correlated below (respectively, above) the threshold of 0.
See Methods for information on how bivariate noise was generated with
asymmetric tail associations.
\label{fig_intro_cop}}
\end{center}
\end{figure*}
<!--***DAN: Shyamolina, it occured to be there are simple and good analytic reasons
why ktau_common is exactly 0.75 (up to sample variation). 
I bet you will see them too if you look up the
definition of kendall tau and think about the extreme tail dependent cases. 
Currently, around line 50 of the sup mat Rmd, you calculate ktau_common
by simulating 100000 points from the extreme-tail-denepdnece noise and then
calculating the sample kendall. Instead, can you please replace those lines 
with the simple line "ktau_common<-0.75"? I would do this myself but I
don't want to risk messing up your code in some way I don't understand. You should not
need the seed anymore to ensure you get the same ktau_common on each run
since you won't be using the sample correlartion anymore to estimate ktau_common,
you'll just be using the analytic result that the population correlation
of data from the extreme-tail-dependent algorithm (in the limit of infinite
sample sizes) is exactly 0.75. Please
let me know if you do not understand something here.-->
<!--***DAN: Shyamolina, currently the sample kendalls are the same on A and D, and
also on B and C, and I know why. This looks weird, and makes the reader
wonder why this is the case. Rather than plotting the reversed version of the
data from A on D, and the reversed version of the data from B on C, can you plase
generate separate, independent data for A and D, and also for B and C? That
will have the effect of making the sample kendalls on the 4 panels slightly
different, which will mean readers think anything funny is going on.-->

@@

<!--***DAN: Shyamolina, the below text makes me wonder if I am misunderstanding something, please
explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
of that and what does it have to do with extinction riks? I would have thought extinction risk
was just the fraction of 10000 sims that went extinct.

@@

<!--***DAN: got to here 2019 10 25-->

We simulated our metapopulation models with noises $(\epsilon_t^i)$ that were 
independent through time with same spatial correlations 
between patches (to within sampling error)
for every simulation, but were moderately (or extremely) left- and 
right-tail associated between patches in
different simulations. For moderately left-tail associated 
noises we drew points from a $n$ dimensional 
Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail association. We generated
extreme left- or right-tail associated noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}. Our models (Table \ref{tab_model_summary}) 
computed extinction risk after $25$ years for $10000$ simulations at $n = 5$ patches
with spatially correlated noises having either moderate or extreme tail 
association in left and right-tail. <!--say how you set ext. threshold = Pe/10-->
We started with initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ for each patch, where 
$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
We assumed the population went extinct after each time step if it was below a 
certain threshold (set to 10\% of the initial population $P_e$ for individual model).

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively. Superscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```

>>>>

<!--concept for correlation in the extreme tail: tail association-->
Though earlier literature considered the influence of spatial synchrony 
in environmental variables on metapopulation extinction risk, to 
our knowledge the distinct influence of "asymmetric tail associations" between
environmental time series in different locations has hardly been considered. 
"Tail associations" are
associations between two random quantities in their extreme values. 
For instance, if two positively associated 
variables are more closely associated in their smaller values than in their larger values,
they are said to have stronger
"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). 
If two positively associated variables are more associated in their larger 
values than in their smaller values, they are 
said to have stronger "right-tail" or "upper-tail" association 
(Figure \ref{fig_intro_cop}C, D). A precise definition of a measure of tail 
association is given in Methods.
Figure \ref{fig_intro_cop} shows that two variables can have a wide 
range of patterns of tail association, from extreme left- to extreme right-tail
association, with the same overall correlation. So tail association
is a distinct concept from correlation itself, and may have distinct ecological
ramifications [@Ghosh_copula]. Asymmetric tail associations have been documented 
between environmental variables, and between the same environmental variable 
measured in different places [@Serinaldi2008; @li2013; @She2018; @Goswami2018]. 
Asymmetric tail associations have also been documented between numerous ecological
variables and between environmental and ecological variables [@Ghosh_copula].

@@

There are also reasons to hypothesize that asymmetry of tail association in 
environmental variables may influence extinction risks for metapopulations.
For instance, assuming for the sake of a simple example that low values of an environmental
variable are "bad" for populations of a given species and high values are "good",
greater left-tail associations between measurements of the variable in different
habitat patches may cause higher metapopulation extinction risk because then
very bad years for component populations would occur at the same time in 
many patches, reducing the potential for rescue effects. Greater right-tail associations
would probably not have the same effect, even if overall correlations between locations
were the same, because in that case it would be very good years for the component 
populations that would occur at the same time in many patches [@Ghosh_copula]. 
@Ghosh_copula substantiated this intuition by showing that it held true for a 
modeling setup using a spatial version of a very simple non-density-dependent
population growth model (the Lewontin-Cohen model). In essence, if environmental
"catastrophes" (i.e., extremely bad years for a population) are widely 
spatially synchronized it should create much greater metapopulation extinction 
risk than if "bonanzas" (i.e., extremely good years) are widely synchronized, even
if overall (non-tail-specific) levels of environmental synchrony are the same
in both these cases. The present study extends tests of this idea to metapopulation
models incorporating density dependence.

@@

<!--objectives, last para of intro-->
In this paper, we will explore the potential for asymmetric tail associations to 
influence extinction risk in populations and metapopulations. Specifically,
we will address the following sets of questions. Q1) Can the nature of tail associations
between vital rates in an age-structured population model influence
extinction risk for the model, and to what extent? Q2) Does the nature
of tail associations between measurements of an environmental variable in different
locations influence the extinction risk of a density-dependent metapopulation model
that depends on the environmental variable? How does the nature of the density
dependence mediate this influence? Will density dependence reverse or reproduce
the earlier result [@Ghosh_copula], using non-density-dependent models, that 
left-tail dependence in environmental variables accentuates and right-tail dependence 
mitigates extinction risks?
For Q1, we simulate a simple 2-stage model. We hypothesize that left-tail dependence
between juvenile survival and adult fecundity rates increases extinction risk
relative to right-tail dependence, even when the overall correlation is kept fixed. 
For Q2, we simulate spatial extensions of several simple and commonly
used population models. We intentionally use a simulation approach and simple models
instead of attempting mathematical analysis and/or using models that are more realistic, 
because our intention here is to provide an initial exploration only of 
the potential mechanisms outlined above which suggest that
asymmetries of tail association may make a difference for extinction risks. 
Some material in the Discussion outlines possible next steps.
<!--***DAN: In Discussion need to talk about what analytic results are needed
to take the next step of determining when tail depencence may make enough of
a difference to think about. I think our results, together with Tulja's 
approximation, suggest that tail associations probably don't make much difference
for models without a spatial component. But our results illsutaret that it can
make a big difference when there is a spatial component, apparently because of 
the influence of tail associations on rescue effects. Could maybe even outline
analytic approaches for non-dd models? Not sure what would be done with dd
models. The point is you need to take the next step in the Discussion of 
talking about how our results, which admitedly are just for specific models,
could be turned into analytic results or general results. Interestingly, 
Tulja's approximation may also apply to the spatial case, at least when 
all locations can be reached from all other locations via dispersal. But the 
approximation seems poor in that case. Need to think about it more, and maybe
need more research.-->
Though tail associations are used
to assess risks in fields such as finance and hydrology 
[@alexander2008market; @li2013; @chen2011flood; @Goswami2018], to our
knowledge asymmetric tail associations have rarely been assessed for their
importance for extinction risks of populations and species (but see @Ghosh_copula). 
Our questions are an important step for understanding
the potential influence of ECEs on extinction risks if ECEs are associated with
tail dependence in environmental variables.

@@

<!--FIG : pedagogical fig showing 4 types of tail associations: 
extreme left, moderate left, moderate right, extreme right-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
\caption{Two stochastic quantities can have same Kendall 
correlation, but different degrees of tail association: 
(A) extreme left-tail association, (B) moderate left-tail association, (C) 
moderate right-tail association, and (D) extreme right-tail association.
In all cases, Kendall's $\tau$ was the same, $\tau = 0.75$,
up to sample variation (the sample correlation for each dataset is displayed
on its panel). The extreme left-tail (respectively, right-tail) dependent 
case is perfectly correlated below (respectively, above) the threshold of 0.
See Methods for information on how bivariate noise was generated with
asymmetric tail associations.
\label{fig_intro_cop}}
\end{center}
\end{figure*}
<!--***DAN: Shyamolina, it occured to be there are simple and good analytic reasons
why ktau_common is exactly 0.75 (up to sample variation). 
I bet you will see them too if you look up the
definition of kendall tau and think about the extreme tail dependent cases. 
Currently, around line 50 of the sup mat Rmd, you calculate ktau_common
by simulating 100000 points from the extreme-tail-denepdnece noise and then
calculating the sample kendall. Instead, can you please replace those lines 
with the simple line "ktau_common<-0.75"? I would do this myself but I
don't want to risk messing up your code in some way I don't understand. You should not
need the seed anymore to ensure you get the same ktau_common on each run
since you won't be using the sample correlartion anymore to estimate ktau_common,
you'll just be using the analytic result that the population correlation
of data from the extreme-tail-dependent algorithm (in the limit of infinite
sample sizes) is exactly 0.75. Please
let me know if you do not understand something here.-->
<!--***DAN: Shyamolina, currently the sample kendalls are the same on A and D, and
also on B and C, and I know why. This looks weird, and makes the reader
wonder why this is the case. Rather than plotting the reversed version of the
data from A on D, and the reversed version of the data from B on C, can you plase
generate separate, independent data for A and D, and also for B and C? That
will have the effect of making the sample kendalls on the 4 panels slightly
different, which will mean readers wont think anything funny is going on.-->

@@

<!--***DAN: Insert para here talking about in what sense the variables
$s_e(t)$ and $f_a(t)$ are asymmetrically associated, through time, in their 
tails. This is where you intro the partial Spearman and may also make references
to Genest and others indicating that, when you are interested in dependence,
you want a rank-based notion of association - it's the only thing that
separates the association from the marginals. I will insert some text now,
but it won't fit perfectly until after we go through and replace Kendall with
Spearman throughout.-->
The variables $u(t)$ and $v(t)$ are visibly asymmetrically tail associated (Fig. \ref{fig_intro_cop}),
but we have instead used the transformed variables $s_e(t)$ and $\tilde{f}_a(t)$
in our model. In what sense are $s_e(t)$ and $\tilde{f}_a(t)$ asymmetrically
tail associated? Given two boundary quantities $l_b$ and $u_b$ such
that $0 \leq l_b < u_b \leq 1$, and letting $x(t)$ be the rank of 
$u(t)$ in the set $\{u(t) | t=1,\ldots,T\}$ and letting 
$y(t)$ be the rank of 
$v(t)$ in the set $\{v(t) | i=1,\ldots,T\}$ (smallest elements
are defined to have rank 1),
consider the portions of the 
lines $x+y = 2l_b$ and $x+y=2u_b$ which meet with the unit square
(Fig. \ref{PartialSpearmanFig}). 
<!--***DAN: need to make this figure-->
We can then define the *partial Spearman correlation*,
\begin{equation}
\cor_{l_b,u_b}(u,v) = \frac{\sum (x(t)-\mean(x))(y(t)-\mean(y))}{(n-1)\sqrt{\var(x) \var(y)}},\label{PartialSpearmanEq}
\end{equation}
\noindent where sample means and variances in this equation
are computed using all the data but the sum is computed only
over the $t$ such that $x(t)+y(t) > 2l_b$ and $x(t)+y(t)<2u_b$. 
The partial Spearman correlation is the component of the Spearman
correlation of $u$ and $v$ which can be attributed to the points
for which the ranks, $x$ and $y$, lie between the boundary lines
we defined (Fig. \ref{PartialSpearmanFig}) in
the unit square. The quantity $\cor_{0,0.5}(u,v)-\cor_{0.5,1}(u,v)$ 
is a way to measure asymmetry of tail associations between
$u$ and $v$. And this quantity is the same as 
$\cor_{0,0.5}(s_e,\tilde{f}_a)-\cor_{0.5,1}(s_e,\tilde{f}_a)$, 
because the partial Spearman correlation, like the Spearman
correlation itself, is based on ranks. Ranks are the same
for $u(t)$ and for $s_e(t)$ because these quantities are 
related by a monotonic transformation (and likewise for 
$v(t)$ and $\tilde{f}_a(t)$). Thus, in the sense captured
by the statistic $\cor_{0,0.5}-\cor_{0.5,1}$, $s_e(t)$
and $\tilde{f}_a(t)$ have the same asymmetry of tail
association as the values $u(t)$ and $v(t)$ from which they
were constructed.

<!--***DAN: Shyamolina, the below commented text makes me wonder if I am misunderstanding something, please
explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
of that and what does it have to do with extinction riks? I would have thought extinction risk
was just the fraction of 10000 sims that went extinct.

@@

We simulated our metapopulation models using environmental noises $\epsilon_i(t)$ (Table \ref{tab_model_summary}), 
independent through time, that were generated to have the same overall associations 
between all patches and in all simulations (Kendall's $\tau$ again $0.75$, up to sampling variation), 
but with different degrees and types of asymmetry of tail associations 
between patches in different simulations, as follows. For a simulation of length $T$,
we used methods that can generate $T$ independent $n$-tuples $(u_1(t),\ldots,u_n(t))$
for $t=1,\ldots,T$ with $\tau=0.75$ between $u_i(t)$ and $u_j(t)$ for all $i \neq j$
(up to sampling variation), but with either extreme left-, moderate left-, moderate right-,
or extreme right-tail association between all pairs of locations (as in Fig. \ref{fig_intro_cop}). 
Pairs $u_i(t)$ and $u_j(t)$ looked like Fig. \ref{fig_intro_cop} and were structured, statistically,
in the same way as the stochastic quantities $u(t)$ and $v(t)$ used for the non-spatial 
simulations described above. For any $i$, $u_i(t)$ was again uniformly distributed on 
the interval $(0,1)$. To obtain the values $\epsilon_i(t)$, the $u_i(t)$ were 
transformed so that the resulting numbers were
normally distributed with mean $0$ and standard deviation $\sigma$ (Table \ref{tab_model_summary}), 
for all $i$. This was 
accomplished with the inverse cummulative distribution function of the normal distribution
with mean $0$ and standard deviation $\sigma$. Because this was a monotonic transformation,
the Kendall correlation between $\epsilon_i(t)$ and $\epsilon_j(t)$ was the same 
as that between $u_i(t)$ and $u_j(t)$. The quantities $(u_1(t),\ldots,u_n(t))$
for $t=1,\ldots,T$ were generated as described in Appendix \ref{SM-makenoise}.

<!--***DAN: Dan is starting to wonder if referees will wonder something like "what 
happens to tail associations when you do these transformatons." We do not wonder 
this because we have in mind a defn for tail associations that only depends on ranks,
but we have not presented that defn. We have instead only given an intuitive description
of tail association. This may be a problem. Referees will see uf generating data with tail associatons
and uniform marginals, but then transforming it, and putting the transformed values into 
our models. And we are confortable with the idea that that transformation does not
alter the tail associations, but they are not comfortable with that.
We may need to define tail associations
using partial Spearman or something, to get this idea across. The problem with that is, 
we use Kendall throughout
the rest of the ms, and so if we define partial Spearman that will create an inconsistency
(using Kendall in some places, Spearman in others) that may raise eyebrows in a short paper 
like this. Not sure right now what to do anout this, we need to think about it. Can we add a new
para at the beginnign of methods giving some kind of formal introduction to tail associations
that makes clear it is something we consider to be for ranks? Without using partial Spearman?
There is no partial Kendall, since Kendall is based on concordant and discordant pairs.
An alterntive is to replace Kendall wth Spearman throughout the paper, then go ahead
and intro partial Spearmn as a way of making it clear tail associaton is for ranks. 
Is this possible? It would definitely be a pain, so therefore want to avoid unless
necessary. One thing this would require would be calculating the populaton (not sample) 
Spearman of the extreme tail associated data. At least we could estimate it very precisely
using a huge sample (would not have the same problems doing that as we did for Kendall),
and may even be able to calculate it analytically. Let's just use Spearman and calculate it very
precisely using a large sample. Well, maybe we could calculate it analytically. 
-->

To calculate extinction risk for a given model and a given choice of tail association,
we simulated the model for $25$ time steps, $10000$ times, using $n=5$ habitat patches.
The initial condition $P_e$ was used in each habitat patch ($P_e$ is the equilibrium value of the 
one-patch, deterministic version of the model, see Table \ref{tab_model_summary}).
After each time step, if $P_i(t)$ was less than $P_e/10$ then it was set to $0$.
The whole metapopulation was considered to have gone extinct if all patches were $0$
a time step $25$.
<!--***DAN: Shyamolina, please check the above two sentences, is this what you did?-->
Extinction risk was the fraction of 10000 metapopulations that went extinct.

<!--***DAN: Dan plans to delete this text as it has been superseded by the above. Keeping it
around until I am sure the above is right:
Our models (Table \ref{tab_model_summary}) 
computed extinction risk after $25$ years for $10000$ simulations at $n = 5$ patches
with spatially correlated noises having either moderate or extreme tail 
association in left and right-tail. 
We started with initial population $P_{t=0}^i = P_e^i, (i = 1,2,\ldots,n)$ for each patch, where 
$P_e^i$ was the equilibrium value (or similar to carrying capacity concept) 
for each density-dependence model (see 4$^{th}$ column of Table \ref{tab_model_summary}).
We assumed the population went extinct after each time step if it was below a 
certain threshold (set to 10\% of the initial population $P_e$ for individual model). 
-->

<!--***DAN: Dan stopped here on 2019 10 27 until we resolve some of the issues that have come up-->

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of six density dependence metapopulation models considered for each patch; last three columns indicate equilibrium population value ($P_e^{i}$), critical growth rate ($r_c^{i}$ : transition point from under- to over-compensatory dynamics of deterministic models) and bifurcation parameter ($r_{bf}^{i}$ : transition point from stable to unstable equilibrium of deterministic models), respectively. Superscript $i=1,2,...,n$ indicates index for $i^{th}$ patch.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```
<!--***DAN: Shyamolina, in this table and in the sup mat (and anywhere else where this comment applies),
please replace the notation P^i_t and \epsilon^i_t by P_i(t) and \epsilon_i(t), the reason being,
putting an i in the superscript makes it look like and exponent. This is legitimately confusing because
we use lots of other exponents. I want you to generally refrain from placing the abitat patch index in the
superscript, putting in the subscript instead. This means you have to put the t in parens in front of the
variable in question, instead of in the subscript. This will require several changes to this table. Please
make those changes and then commit. Then please make the corresponding changes to the sup mat, whereever needed.-->
<!--***DAN: Dan, once Shyamolina does the above, you will then have to go back and edit this table in various
ways that you cannot do until after that change is made.-->

### 2019-10-28 SuppMat.Rmd
working on para intro-ing partial spearman

>>>>

<!--yet more stuff from the main text-->

For moderately left-tail associated 
noises we drew points from a $n$ dimensional 
Clayton copula with same Kendall correlation ($\tau =$ `r ktau_common`) between
any two out of $n$ patches and made 
the marginals normal with zero mean and standard deviation $\sigma$ 
(see the specified value $\sigma$ for each model in 3$^{rd}$ 
column of Table \ref{tab_model_summary}). For moderately 
right-tail associated noises we flipped ($180^{\circ}$ rotation) the points drawn 
from Clayton copula so that all the statistical properties remained 
same except for the direction of tail association. We generated
extreme left- or right-tail associated noises via algorithm stated in 
Appendix \ref{SM-extr_tail_dep}.
